---
title:  "Group Normalization - Data Acquisition" 
author: "[Francois Collin](https://www.linkedin.com/in/francoisz/)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
always_allow_html: yes
output:
  # for gtanse of .tabset
  #bookdown::html_document2:
  html_document:
    code_folding: hide
    code_download: true
    toc: true
    toc_depth: 2
    # does this have an effect
    fig_caption: yes
    # this has no effect
    number_sections: yes
    # css: ['../_css/pandoc3.css', '../_css/myMargins.css']
bibliography: [../Refs/group_norm.bib,../../_bibFiles/_healthy_aging.bib, ../../_bibFiles/_Breiman.bib, ../../_bibFiles/_Freedman.bib, ../../_bibFiles/_Yu.bib, ../../_bibFiles/_RUV.bib, ../../_bibFiles/_RMA.bib, ../../_bibFiles/_scRNAseq_norm.bib] 
csl: ../../_csl/cell-numeric.csl
link-citations: true 
---

`r LONG_EVAL = T`

<!--
### LONG_EVAL = `r LONG_EVAL`  {-}
-->

```{r get-data-Prelims,  include=FALSE, echo=FALSE, results='hide', message=FALSE} 

FN <- "_get_data" 
if(sum(grepl(FN, list.files()))==0) stop("Check FN")

PREFIX <- FN

 suppressMessages(require(rmarkdown))
 suppressMessages(require(knitr))

 suppressPackageStartupMessages(require(methods))
 suppressPackageStartupMessages(require(bookdown))

 suppressPackageStartupMessages(require(magrittr))

 # Shotcuts for knitting and rendering while in R session (Invoke interactive R from R/Scripts folder)
 kk <- function(n='') knitr::knit2html(paste("t", n, sep=''), envir=globalenv(),
       output=paste(FN,".html", sep=''))

 rr <- function(n='') rmarkdown::render(paste("t", n, sep=''), envir=globalenv(),
       output_file=paste(FN,".html", sep='')) ##, output_dir='Scripts')
 bb <- function(n='') browseURL(paste(FN,".html", sep='')) 
 # Theual shortcuts 
 zz <- function(n='') source(paste("t", n, sep='')) 

 WRKDIR <- '..'
 if(!file.exists(WRKDIR)) stop("WRKDIR ERROR", WRKDIR)


 # file rmarkdown file management options: cache, figures
 cache_DIR <- paste0(PREFIX, '_cache/')
 suppressMessages(dir.create(cache_DIR, recursive=T))
 opts_chunk$set(cache.path=cache_DIR)

 # NOTE: need to add PREFIX to figure path
 # due to confluence "feature" of keeping the filename of embedded figures

 figure_DIR <- paste(PREFIX, '_figures')
 suppressPackageStartupMessages(dir.create(figure_DIR, recursive=T))
 opts_chunk$set(fig.path=paste0(figure_DIR, PREFIX))

 # need a local copy of help_DIR
 #help_DIR <- file.path(WRKDIR,'Scripts', 'help_files')
 help_DIR <- file.path('.', 'help_files')
 suppressMessages(dir.create(help_DIR, recursive=T))
   
 temp_DIR <- file.path('.','Scripts', 'temp_files')
 suppressMessages(dir.create(temp_DIR, recursive=T))


``` 
<!-- ######################################################################## -->

***

```{r get-data-utilityFns, echo=FALSE}
 # Here we define some utility functions
source('r/utilityFns.r')

```


<!--

* What data?  Where the data are.  Where they are not.  Etc.

-->

<br/>

## Run Parameters {-}

```{r get-data-params, cache=T, cache.vars='params_lst'}

params_lst <- list( 
   AE_ID  = "E-MEXP-1172", 
   files_loc = "../extData/E-MEXP-1172", 
   platform_ID = "A-AFFY-42", 
   designTable = "A-AFFY-42.additional.txt")

if(F) 
data.frame( 
   feature = names(params_lst), 
   value = unlist(params_lst)) %>%
knitr::kable( 
row.names = F, 
caption = "Run Parameters")

```

```{r get-data-params-2, results = 'asis'} 
   cat(sprintf( "- <span style = 'color:grey'>%s:</span> `%s`  \n", 
   names(params_lst), params_lst))  

```

<br/>

# Array Design {.tabset}

## Home {-}

```{r get-data-read-desiqn, cache = T, cache.vars = "AE_array_frm", eval=F}

AE_array_frm <- with(params_lst, 
data.table::fread(file.path("../extData", platform_ID, designTable)))

AE_array_frm %<>% dplyr::select(-V8) %>% dplyr::rename(regionName = targetName)

saveObj(paste0("_get_data_",params_lst$platform_ID, "_array_frm"),
"AE_array_frm")
``` 
```{r} 
loadObj(
  paste0("_get_data_",params_lst$platform_ID, "_array_frm"), 
  "AE_array_frm")
```

Define functions ...

```{r get-data-def-fun, cache=T, cache.vars = c("find_lcs_str_f")}

# Define function to find the longest common substring.

# A common approach is to find the actual substring with a custom function.
# The following Stack Overflow example uses a dynamic programming approach to
# find the longest common substring.  Credit:
# https://stackoverflow.com/a/35384317/989255

# Troubleshooting
xxx  <- function() { 
  JJ <- 1 
  a <- AE_array_plus_frm$reporterRefSeq[JJ] 
  b <- AE_array_plus_frm$reporterRefSeq[JJ+1] 
}


find_lcs_str_f <- function(a, b) { 
  Result <- c( 
  loc = NA, 
  len = NA, 
  lcs = NA)

  if(!is.na(a) & !is.na(b)) { 
    A <- strsplit(a, "")[[1]]         # from strib 
    B <- strsplit(b, "")[[1]] 
    L <- matrix(0, length(A), length(B)) 
    ones <- which(outer(A, B, "=="), arr.ind = TRUE) 
    ones <- ones[order(ones[, 1]), ]
  
    for (i in 1:nrow(ones)) { 
      v <- ones[i, , drop = FALSE] 
      L[v] <- ifelse(any(v == 1), 1, L[v - 1] + 1) 
    }
  
    max_len <- max(L) 
    if (max_len == 0) return("")
    
    max_loc <- which(L == max_len, arr.ind = TRUE)[1, ] 
    max_str <- paste0(A[(max_loc[1] - max_len + 1):max_loc[1]], collapse = "")
  
    # Changed from  list()
    Result = c( 
    loc = max_loc[[1]] - max_len[1] + 1, 
    len = max_len, 
    lcs = max_str) 
   } #if(!is.na(a) & !is.na(b))
   return(Result) 
} #find_lcs_str_f

```


## Overview ## {-}

<!--

<p><p/>
* Read and save table as is.
   - preserve the ordering as it will be assumed that the cel files are stored
     in the same order 
-->


<p><p/>
* The array has 2560 x 2560 = 6553600 cels which are described
by 7 descriptors in  `r params_lst$designTable`.  
   - The first 4 columns describe the spots on the array:
      - **Row, Column, reporterName, reporterSequence** 
   - The last 3 columns describe the targeted sequence in yeast:
      - **regionName (yeast chrom), targetStart, targetEnd**


<p><p/>
* Of the `r nrow(AE_array_frm)` cells.
   - `r sum(!is.na(AE_array_frm$regionName))` have non-missing yeast genome
annotation in `r params_lst$designTable`.
   - `r round(100*sum(!is.na(AE_array_frm$regionName))/nrow(AE_array_frm))`% of
     array spots or cells.
   - the rest of the cels are part of tthe annotation on this file.

params_lst <- list( 
   AE_ID  = "E-MEXP-1172", 
   files_loc = "../extData/E-MEXP-1172", 
   platform_ID = "A-AFFY-42", 
   designTable = "A-AFFY-42.additional.txt")

if(F) 
data.frame( 
   feature = names(params_lst), 
   value = unlist(params_lst)) %>%
knitr::kable( 
row.names = F, 
caption = "Run Parameters")

```

```{r get-data-params-2, results = 'asis'} 
   cat(sprintf( "- <span style = 'color:grey'>%s:</span> `%s`  \n", 
   names(params_lst), params_lst))  

```

<br/>

# Array Design {.tabset}

## Home {-}

```{r get-data-read-desiqn, cache = T, cache.vars = "AE_array_frm", eval=F}

AE_array_frm <- with(params_lst, 
data.table::fread(file.path("../extData", platform_ID, designTable)))

AE_array_frm %<>% dplyr::select(-V8) %>% dplyr::rename(regionName = targetName)

saveObj(paste0("_get_data_",params_lst$platform_ID, "_array_frm"),
"AE_array_frm")
``` 
```{r} 
loadObj(
  paste0("_get_data_",params_lst$platform_ID, "_array_frm"), 
  "AE_array_frm")
```

Define functions ...

```{r get-data-def-fun, cache=T, cache.vars = c("find_lcs_str_f")}

# Define function to find the longest common substring.

# A common approach is to find the actual substring with a custom function.
# The following Stack Overflow example uses a dynamic programming approach to
# find the longest common substring.  Credit:
# https://stackoverflow.com/a/35384317/989255

# Troubleshooting
xxx  <- function() { 
  JJ <- 1 
  a <- AE_array_plus_frm$reporterRefSeq[JJ] 
  b <- AE_array_plus_frm$reporterRefSeq[JJ+1] 
}


find_lcs_str_f <- function(a, b) { 
  Result <- c( 
  loc = NA, 
  len = NA, 
  lcs = NA)

  if(!is.na(a) & !is.na(b)) { 
    A <- strsplit(a, "")[[1]]         # from strib 
    B <- strsplit(b, "")[[1]] 
    L <- matrix(0, length(A), length(B)) 
    ones <- which(outer(A, B, "=="), arr.ind = TRUE) 
    ones <- ones[order(ones[, 1]), ]
  
    for (i in 1:nrow(ones)) { 
      v <- ones[i, , drop = FALSE] 
      L[v] <- ifelse(any(v == 1), 1, L[v - 1] + 1) 
    }
  
    max_len <- max(L) 
    if (max_len == 0) return("")
    
    max_loc <- which(L == max_len, arr.ind = TRUE)[1, ] 
    max_str <- paste0(A[(max_loc[1] - max_len + 1):max_loc[1]], collapse = "")
  
    # Changed from  list()
    Result = c( 
    loc = max_loc[[1]] - max_len[1] + 1, 
    len = max_len, 
    lcs = max_str) 
   } #if(!is.na(a) & !is.na(b))
   return(Result) 
} #find_lcs_str_f

```


## Overview ## {-}

<!--

<p><p/>
* Read and save table as is.
   - preserve the ordering as it will be assumed that the cel files are stored
     in the same order 
-->


<p><p/>
* The array has 2560 x 2560 = 6553600 cels which are described
by 7 descriptors in  `r params_lst$designTable`.  
   - The first 4 columns describe the spots on the array:
      - **Row, Column, reporterName, reporterSequence** 
   - The last 3 columns describe the targeted sequence in yeast:
      - **regionName (yeast chrom), targetStart, targetEnd**


<p><p/>
* Of the `r nrow(AE_array_frm)` cells.
   - `r sum(!is.na(AE_array_frm$regionName))` have non-missing yeast genome
annotation in 
   - `r round(100*sum(!is.na(AE_array_frm$regionName))/nrow(AE_array_frm))`% of
     array spots or cells.
   - the rest of the cels are part of tthe annotation on this file.

<br/>

## Content Design Highlights {-}


<p><p/>
* See <span style="font-variant:small-caps;">methods}</span> section
in Lee et al. (2007) [@Lee:2007aa].

<p><p/>
* Interrogates the **<span style = 'color:blue'>entire genomic  sequence</span>**,
including repetitive elements. 
   - the repetitive elements provide no useful and may have been filtered out of
the annotation file.
      - What query can be run to verify this?

<p><p/>
* Includes oligonucleotides to interrogate the 
<span style = 'color:blue'>**forward** and **reverse** strands</span>. 

* **<span style = 'color:blue'>25mer probes</span>** were tiled 
at **<span style = 'color:blue'>8bp intervals</span>** on each strand^[
measured from the central position of each oligo] creating a
sense specific overlap of 17bp. 

<p><p/>
* The oligos detecting **<span style = 'color:blue'>sense and 
antisense targets were offset by 4bp</span>** to yield  a high resolution view
of the genome
   - For a fixed region, the  capture  of targets from the complimentary strands
     involves the same diploid/double stranded sequence, but the sequence bound to
the array differs.
      - when we combine counts across the two strand specific capture modes and
compare across samples, we will see strand effects in many cases.

<!-- May be incorrect - see _get_data_truncated_pobes.md
<p><p/>
* **approximately <span style = 'color:blue'>2% of total probes were 
truncated</span>** to economize on the required number of oligonucleotide
synthesis steps.
   - this is not reflected in the probe information that we have.
-->

<br/>

##  Design Verification {- .tabset}

### Home {-}

<br/>

### Information {-}

<p><p/>
* All of our design information comes from `r params_lst$designTable`, which
was downloaded from **ArrayExpress#E-MEXP-1172**.
   - the dataset contains probe and target (yeast region) information for 41%
     of the array loci or cells.
   - only yeast targets are annotated ( ~ 41% of cels on chip)
   - we don't know that *all* yeast regions targeted by the array are captured.
      - ie, some filtering may have been done which would explain some of the
        gaps in the coverage indicated by the dataset.


<!--
<p><p/>
* 2% of probes should be truncated (source??)
   - truncation of probes mentioned in the array description is not captured in
     the array design file/doc - all probes are annotated as 25-mers.
-->

<br/>

### Preparation {-}


<p><p/>
* Order targets by Reference Position, compute inter-target offset:
should be 4 bp shifts between alternating sense strands
   - **Reference Position** for probes is the **Start Position** on the sense
     strand, and the **End Position** on the anti-sense strand.
      - **sense strand probes** are depicted on the sense strand coordinate axis in
        the obvious way - line up the start of the probe to the axis
coordinate.
      - for the **anti-sense probes**, the anti-sense strand coordinates are put
        down in the reverse direction - the last anti-sense strand coordinate
aligns with the first sense strand coordinate, and the first anti-sense strand
coordinate aligns with the last sense strand coordinate.
      - **a reference position occurs every 4bp on alternating strands**.
when the start position of sense strand probes are at 8 bp intervals,
and the end position of anti-sense strand probes are at 8 bp intervals
shifted 4bp from the sense strand probe position, 
<!--
      - an alternative design would be to lay the anti-sense probes at the same
        loci as the sense strand probes; 0 bp shift.
	 - **under what assumptions is each design preferable.?**
	    - only if strand effects are of interest 
-->

<!--
* Verify matching sequence in anti sense probes
-->

<p><p/>
* With the targets ordered by reference position within region, the target layout
can be compared to the design by examining the following quantities:
<p><p/>
   - **sense2 = sense * lag(sense)**
      - $\equiv$ -1 as positive (negative) sense
     targets should be followed by a negative (positive) sense targets  4bp
away.
<p><p/>
   - **diffPos = refPos - lag(refPos)** : should be 4 by design
<p><p/>
   - **loc, len** = the starting position and length of the longest sequence in the
     current reporter sequence also found in previous reporter sequence
       - should be 5 and 21, respectively, by design.

<p><p/> 
* Let the quartet **Sign = (sense2, diffPos, loc, len)** denote a target's signature or
phenotype.  Signatures of interest include:
   - **(-1, 4, 5, 21)** - the Design signature
   - **(+/- 1, 0, 1, 25)** -  Replicate target signatuge (exact same sequence)

<p><p/> 
* Other interesting target sequences are indicated by runs sense2 == 1, indicating
that only one strand is interrogated for a stretch of the genome.

<br/>

```{r get-data-by-target, cache=T, cache.vars=c("AE_array_plus_frm"), eval=T & LONG_EVAL}

AE_array_plus_frm <- AE_array_frm %>% 
dplyr::mutate(
sense = sign(targetEnd - targetStart), 
targetPos = ifelse(sense == 1, targetStart, targetEnd), 
rcSeq = toupper(spgs::reverseComplement(reporterSequence)), 
reporterRefSeq = ifelse(sense == 1, reporterSequence, rcSeq),
                
probeSize = nchar(reporterRefSeq), 
targetLength = abs(targetEnd - targetStart) + 1
) %>% 
dplyr::arrange(regionName, targetPos) %>%
dplyr::filter(!is.na(regionName))

  ### DONT SAVE AS NAME COLLIDES WITH ENHANCED VERSION

```




```{r get-data-by-target-2, cache=T, cache.vars=c("AE_array_plus_tib1"), eval=T & LONG_EVAL}
AE_array_plus_tibl <- AE_array_plus_frm %>% 
dplyr::select(-c(probeSize, targetLength, reporterSequence, rcSeq)) %>% 
dplyr::group_by(regionName) %>%
dplyr::arrange(regionName, targetPos) %>% 
dplyr::mutate(
  sense2 = sense * dplyr::lag(sense), 
  strand = ifelse(sense == 1, 'p', 'm'),
  lag_strand = ifelse(dplyr::lag(sense) == 1, 'p', 'm'),
  diffPos = targetPos - dplyr::lag(targetPos), 
  ###regionNum = rank(targetPos),  - NOT NEEDED 
  lcs_result = purrr::map2(
    dplyr::lag(reporterRefSeq),
    reporterRefSeq, 
    find_lcs_str_f)) %>% 
#tidyr::unnest(lcs_result) %>%
tidyr::unnest_wider(lcs_result) %>% dplyr::ungroup()%>%
dplyr::arrange(regionName, targetPos, sense)

AE_array_plus_tibl %<>%
tidyr::unite(
 col = "strand2",
 lag_strand, strand,
 sep = '')

AE_array_plus_tibl %<>% 
dplyr::relocate(regionName, targetPos, strand2,
                diffPos, loc, len, sense, reporterRefSeq, Row, Column, reporterName)

saveObj("_get_data_AE_array_plus_tibl", 'AE_array_plus_tibl')

``` 
```{r} 
loadObj("_get_data_AE_array_plus_tibl", 'AE_array_plus_tibl') 
```

```{r get-data-subset-array, cache=T, cache.vars="AE_array_plus_n100_frm"}

row_lst <- split(rownames(AE_array_plus_tibl),AE_array_plus_tibl$regionName)

AE_array_plus_n100_frm <- do.call("rbind", lapply(row_lst, function(ROWS)
AE_array_plus_tibl[ROWS[1:100], ]))

```



<br/>

### Sample Probe Data {-}

```{r get-data-display-array, fig.cap="Detailed Probe Description - First 100 by Group"}

AE_array_plus_n100_frm %>% DT::datatable(
  extensions = 'Buttons', 
  options = list(pageLength=50), 
  caption = "Detailed Probe Description - First 100 by Group", 
  rownames=F)

```


<br/>

### Probe Data Summaries {- .tabset}

* Probe/Target  signature:
   - prop by design
   - replicates, by strand
   - strand specific runs


<!--
* Only the first region, NC_001133 or chrom 1, have comparable
number of probes.  
-->

<br/>

<p><p/> 
* Let the quartet (strand2, diffPos, loc, len) denote a target's signature or
* phenotype.  Signatures of interest:
   - (mp|pm , 4, 5, 21) - the design signature
   - (mp|pm,  0, 1, 25) - replicate signature (exact same sequence),

<p><p/> 
* Other interesting target sequences are indicated by runs of
strand2 == 'pp' or 'mm' indicating that only one strand is interrogated
for a stretch of the genome.

```{r get-data-probe-sum, cache=T, cache.vars=c("AE_array_plus_tibl", "probe_sign_count_frm")}

AE_array_plus_tibl %<>%
tidyr::unite(
 col = "probe_sign",
 diffPos, loc, len, strand2,
 sep = '~')

probe_sign_count_frm <- AE_array_plus_tibl %>%
dplyr::group_by(probe_sign) %>%
dplyr::summarize(count = dplyr::n()) %>%
as.data.frame()

probe_sign_count_frm %>% 
dplyr::arrange(desc(count)) %>%
head(n=200)


```

<br/>
<br/>

<!-- ########################################### -->

# Read the Cel files

* To read cel files when there is no cdf
`affyio::read.celfile` can be used.

<br/> 


# References {#references}
<div id="refs"></div>
    
<br/> 

```{r , echo=FALSE}
  knit_exit()
```

pander::pander(sessionInfo())
```

  * WRKDIR = `r normalizePath(WRKDIR)`
  * FN = `r FN`
  * RUN DATE = `r date()`


############################################################
## ARCHIVED CODE BELOW {-}
############################################################
# Appendix

## Read Affymetrix (Affy) CEL files {-}

To read Affymetrix (Affy) CEL files from ArrayExpress, you will use the statistical programming language R and packages from the Bioconductor project. The primary Bioconductor packages for this task are for downloading the data and or for reading and preprocessing the CEL files. [1, 2, 3]

Step 1: Install R and Bioconductor
If you don't have it, install R from the CRAN website.
Open R or RStudio and install Bioconductor and the necessary packages by running the following commands: [4]

Step 2: Download the CEL files

You will need the ArrayExpress accession code for the dataset you wish to download (e.g., E-MTAB-2967).
Use the function from the package to automatically download the raw CEL files and experiment metadata.
First, create a directory to store your files. [5]

Step 3: Read and preprocess the CEL files

After downloading the files, you can use either the or package to read and preprocess the data. The package is a more modern alternative suitable for most newer arrays, while is a classic option. [1, 3, 6]

Option A: Using the package
The package is generally recommended for newer Affymetrix arrays and provides robust preprocessing methods.
Load the package.
List the CEL files in your download directory.
Read the CEL files into an object.
Perform Robust Multi-array Average (RMA) normalization to get gene expression values.
Access the expression matrix.

Option B: Using the package
The package is a classic tool for handling Affymetrix data.
Load the package.
Set your working directory to the folder containing your CEL files.
Read the CEL files into an object.
Perform RMA normalization to create an object.
Access the expression matrix. [3]

Step 4: Further analysis

After obtaining the expression matrix, you can proceed with your analysis, such as differential gene expression, clustering, or other bioinformatics methods. It is often useful to perform quality control checks on the raw data before normalization. [4, 7, 8, 9, 10]


AI responses may include mistakes.
[1] https://gtk-teaching.github.io/Microarrays-R/03-Supplemental-Affy/index.html
[2] https://www.bioconductor.org/packages/release/workflows/vignettes/maEndToEnd/inst/doc/MA-Workflow.html
[3] https://bioconductor.org/packages/devel/bioc/vignettes/affy/inst/doc/affy.pdf
[4] http://bioconductor.jp/packages/3.19/workflows/vignettes/maEndToEnd/inst/doc/MA-Workflow.html
[5] https://pmc.ncbi.nlm.nih.gov/articles/PMC6063319/
[6] https://pmc.ncbi.nlm.nih.gov/articles/PMC10111666/
[7] http://homer.ucsd.edu/homer/basicTutorial/affymetrix.html
[8] https://www.mathematica-journal.com/2013/11/26/detecting-differential-gene-expression-using-affymetrix-microarrays/
[9] https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002276
[10] https://f1000research.com/articles/9-1447



safe_reverseComplement <- function(x, on_error = NA) {
  tryCatch(
    expr = {
      # Attempt the reverseComplement operation
      Biostrings::reverseComplement(x)
    },
    error = function(e) {
      # Handle the error
      warning(paste("An error occurred during reverse complementation:", e$message), call. = FALSE)
      return(on_error)
    }
  )
}

    ###repDNAString =  Biostrings::DNAString(reporterRefSeq),
    ###rcSeq =  Biostrings::reverseComplement(repDNAString)
    ###rcSeq =  safe_reverseComplement(reporterSequence)
## affy::ReadAffy bombs without a cdf.


```{r get-data-read-cel, cache=T, cache.vars="ae_id_affyBatch", eval=F}

ae_id_affyBatch <- affy::ReadAffy(celfile.path = params_lst$files_loc)

```

```
AffyBatch object
size of arrays=2560x2560 features (20 kb)
cdf=Scerevisiae_tlg (??? affyids)
number of samples=6
Error in getCdfInfo(object) : 
  Could not obtain CDF environment, problems encountered:
Specified environment does not contain Scerevisiae_tlg
Library - package scerevisiaetlgcdf not installed
Bioconductor - scerevisiaetlgcdf not available
In addition: Warning message:
missing cdf environment! in show(AffyBatch)

```


   - each target locus on the genome of a sample gives rise 
to two sets of fragments in the extraction, one for each strand.
      - it will be useful to monitor the strand specific differences 
inferred target concentrations as there is built in truth in this context.  
         - there is a 4bp shift between the matching probes


```{r get-dafa-array-data-look, cache=T,cache.vars='',eval = F}

AE_array_frm %>%
#dplyr::filter(!is.na(regionName)) %>%
#dplyr::filter(regionName == "ref|NC_001133|") %>%
dplyr::filter(regionName == "ref|NC_001224|") %>%
dplyr::mutate(orientation = targetEnd - targetStart) %>%
dplyr::arrange(targetPos) %>%
  DT::datatable(extensions = 'Buttons', options = list(pageLength=50),
                caption = "Mito Sequence (001224)")

```




<!-- To run
# nohup Rscript -e "knitr::knit2html('_get_data.Rmd')" > _get_data.log  &

# Or
# nohup Rscript -e "rmarkdown::render('_get_data.Rmd')" > _get_data.log  &

-->
