---
title: "Group Normalization"
###\nLessons from 20 years of expererience"
###author: "[Francois Collin](https://www.linkedin.com/in/francoisz/)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
always_allow_html: yes
output:
  # for gtanse of .tabset
  bookdown::html_document2:
  #html_document:
#   code_folding: hide
#   code_download: true
    toc: true
    toc_depth: 2
    # does this have an effect
    fig_caption: yes
    # this has no effect
    number_sections: yes
    # css: ['../_css/pandoc3.css', '../_css/myMargins.css']
#bibliography: [../Refs/group_norm.bib,../../_bibFiles/_healthy_aging.bib, ../../../_bibFiles/_Breiman.bib, ../../../_bibFiles/_Freedman.bib, ../../../_bibFiles/_Yu.bib, ../../../_bibFiles/_RUV.bib, ../../../_bibFiles/_RMA.bib, ../../../_bibFiles/_scRNAseq_norm.bib]
#csl: ../../_csl/cell-numeric.csl
#link-citations: true
---

```{css sidenote, echo = FALSE}

mmain-container {
    margin-left: 250px;
}
.sidenote, .marginnote { 
  float: right;
  clear: right;
  margin-right: -40%;
  width: 37%;         # best between 50% and 60%
  margin-top: 0;
  margin-bottom: 0;
  font-size: 1.1rem;
  line-height: 1.3;
  vertical-align: baseline;
  position: relative;
  }
```


<style>
@import url('https://fonts.googleapis.com/css?family=Raleway');
@import url('https://fonts.googleapis.com/css?family=Oxygen');
@import url('https://fonts.googleapis.com/css?family=Raleway:bold');
@import url('https://fonts.googleapis.com/css?family=Oxygen:bold');

.main-container {
  max-width: 1400px !important;
}

body{
  font-family: 'Oxygen', sans-serif;
  font-size: 16px;
  line-height: 24px;
}

h1,h2,h3,h4 {
  font-family: 'Raleway', sans-serif;
}

.container { width: 1400px; }

caption {
  font-size: 20px;
  caption-side: top;
  text-indent: 30px;
  background-color: lightgrey;
  color: black;
  margin-top: 5px;
}

g-table-intro h4 {
  text-indent: 0px;
}
</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.width = 8,
                      fig.height = 4)
```

```{r infex-GlobalOptions, results="hide", include=FALSE, cache=FALSE}

knitr::opts_knit$set(stop_on_error = 2L) #really make it stop
options(knitr.table.format = 'html')

options(stringsAsFactors=F)

 #knitr::dep_auto()

```
<!-- ######################################################################## -->


```{r index-Prelims,  include=FALSE, echo=FALSE, results='hide', message=FALSE} 


FN <- "index"
if(sum(grepl(FN, list.files()))==0) stop("Check FN")

PREFIX <- "index" #- replace by FLOWCELL???

 suppressMessages(require(rmarkdown))
 suppressMessages(require(knitr))

 suppressPackageStartupMessages(require(methods))
 suppressPackageStartupMessages(require(bookdown))

 suppressPackageStartupMessages(require(magrittr))

 # Shotcuts for knitting and rendering while in R session (Invoke interactive R from R/Scripts folder)
 kk <- function(n='') knitr::knit2html(paste("t", n, sep=''), envir=globalenv(),
       output=paste(FN,".html", sep=''))

 rr <- function(n='') rmarkdown::render(paste("t", n, sep=''), envir=globalenv(),
       output_file=paste(FN,".html", sep='')) ##, output_dir='Scripts')
 bb <- function(n='') browseURL(paste(FN,".html", sep='')) 
 # The usual shortcuts 
 zz <- function(n='') source(paste("t", n, sep='')) 
``` 

<!-- ######################################################################## -->

<br/>

# Introduction


* We are interested in learning about the preprocessing step referred to as 
normalization in **Group Normalization for Genomic Data**, Ghandi and Beer 2012 [@Ghandi:2012aa]

* David 2006 [@David:2006aa] and Huber 2006 [@Huber:2006aa] offer an alternative  
procedure and provide data and software to assess performance.

<br/>

# AI queries 

* [./claude_3.7_GNGD-chat/](./claude_3.7_GNGD-chat/)

* [./claude_3.7_GNGD-project/](./claude_3.7_GNGD-project/)

* [./claude_3.7_GNGD-project2/](./claude_3.7_GNGD-project2/)

* [./claude_3.7_GNGD-project3/](./claude_3.7_GNGD-project3/)


<br/>

## Data Retrieval

* [_get_data.Rmd](./_get_data.html)


<br/>

# Main Analysis

* [_GN4GD_analysis.html](_GN4GD_analysis.html)





<br/>



# References {#references}
   
<div id="refs"></div>
     
  
 
```{r, echo=FALSE}
  knitr::knit_exit()
```

#############################################################
#############################################################
<!--
* Comparative Illustration of the use of AI Agents 
   -  For Article Review
      - claude (v3.7 , 4.?)
      - chatGPT
      - ???
   - For literature review
-->
#############################################################
#############################################################

