---
title: "Topics on the Statistical Analysis of Genomic Data"
###\nLessons from 20 years of expererience"
###author: "[Francois Collin](https://www.linkedin.com/in/francoisz/)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
always_allow_html: yes
output:
  # for gtanse of .tabset
  bookdown::html_document2:
  #html_document:
#   code_folding: hide
#   code_download: true
    toc: true
    toc_depth: 2
    # does this have an effect
    fig_caption: yes
    # this has no effect
    number_sections: yes
    # css: ['../_css/pandoc3.css', '../_css/myMargins.css']
bibliography: [../Refs/group_norm.bib,../../_bibFiles/_healthy_aging.bib, ../../../_bibFiles/_Breiman.bib, ../../../_bibFiles/_Freedman.bib, ../../../_bibFiles/_Yu.bib, ../../../_bibFiles/_RUV.bib, ../../../_bibFiles/_RMA.bib, ../../../_bibFiles/_scRNAseq_norm.bib]
csl: ../../_csl/cell-numeric.csl
link-citations: true
---

```{css sidenote, echo = FALSE}

.main-container {
    margin-left: 250px;
}
.sidenote, .marginnote { 
  float: right;
  clear: right;
  margin-right: -40%;
  width: 37%;         # best between 50% and 60%
  margin-top: 0;
  margin-bottom: 0;
  font-size: 1.1rem;
  line-height: 1.3;
  vertical-align: baseline;
  position: relative;
  }
```


<style>
@import url('https://fonts.googleapis.com/css?family=Raleway');
@import url('https://fonts.googleapis.com/css?family=Oxygen');
@import url('https://fonts.googleapis.com/css?family=Raleway:bold');
@import url('https://fonts.googleapis.com/css?family=Oxygen:bold');

.main-container {
  max-width: 1400px !important;
}

body{
  font-family: 'Oxygen', sans-serif;
  font-size: 16px;
  line-height: 24px;
}

h1,h2,h3,h4 {
  font-family: 'Raleway', sans-serif;
}

.container { width: 1400px; }

caption {
  font-size: 20px;
  caption-side: top;
  text-indent: 30px;
  background-color: lightgrey;
  color: black;
  margin-top: 5px;
}

g-table-intro h4 {
  text-indent: 0px;
}
</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.width = 8,
                      fig.height = 4)
```

```{r infex-GlobalOptions, results="hide", include=FALSE, cache=FALSE}

knitr::opts_knit$set(stop_on_error = 2L) #really make it stop
options(knitr.table.format = 'html')

options(stringsAsFactors=F)

 #knitr::dep_auto()

```
<!-- ######################################################################## -->


```{r index-Prelims,  include=FALSE, echo=FALSE, results='hide', message=FALSE} 


FN <- "index"
if(sum(grepl(FN, list.files()))==0) stop("Check FN")

PREFIX <- "index" #- replace by FLOWCELL???

 suppressMessages(require(rmarkdown))
 suppressMessages(require(knitr))

 suppressPackageStartupMessages(require(methods))
 suppressPackageStartupMessages(require(bookdown))

 suppressPackageStartupMessages(require(magrittr))

 # Shotcuts for knitting and rendering while in R session (Invoke interactive R from R/Scripts folder)
 kk <- function(n='') knitr::knit2html(paste("t", n, sep=''), envir=globalenv(),
       output=paste(FN,".html", sep=''))

 rr <- function(n='') rmarkdown::render(paste("t", n, sep=''), envir=globalenv(),
       output_file=paste(FN,".html", sep='')) ##, output_dir='Scripts')
 bb <- function(n='') browseURL(paste(FN,".html", sep='')) 
 # The usual shortcuts 
 zz <- function(n='') source(paste("t", n, sep='')) 
``` 

<!-- ######################################################################## -->

<br/>

# Published article review - Ghandi and Beer (2012)

* **Group Normalization for Genomic Data**
by Ghandi and Beer (2012)[@Ghandi:2012ab]) 

* **Main Finding** is that a pre-processing step applied to
Affimetrix tiling array data improves the quality of the data

<p><p/>
* **Nomenclature**.
   - Normalization has different meanings in different context.  
"Normalization" in this paper
refers to a within group or sample probe effect correction.
   - In the context of Omic data analysis, *normalization* is usually
reserved for transformations which affect between sample variability.
      - See [Normalization of omic data after 2007 (INCOB2014)](https://incob.apbionet.org/incob14/assets/INCOB2014/Terry-SpeedInCoB2014.pdf)
and [Identification of Unwanted Variability: Introduction](_intro_IUV.html)
   - A standard language to speak clearly about variability is sorely needed.
      - [claude 3.7 -  Terminology for Variability in Genomic Data Analysis](claude_3.7_GNGD-project/TerminologyVariabillityOmic.html)


<p><p/>
* [Data Analysis Plan](claude_3.7_GNGD-project/GN4GD_implementationGuide.html)

<!--
claude 3.7 reports
   - [Background - Affy Tiling Arrays](claude_3.7_GNGD-chat/affy_tiling_arrays.html0)
-->

<!--  #################################################
## Outline of presentation {-}

<p><p/>
* Describe the Affymetrix Tiling Array design,
the collected data, and typical analyses.
   - we don't have to spend a lot of time on this old technology,
but it's hard to avoid entirely as those data are of the sort
which led to the development of the GN normalization.
   - in the end the question is not so much 
does the preprocessing step work, as **when does it work**.

## Download the data analyzed in the report

* Figures 6 and 7 data are from 
He et al. 2008 [@He:2008aa].
   - [GSE12163](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=%20GSE12163)
* Figure 8 data are from Johnson et al. (2008) [@Johnson:2008aa]
   - [GSE10114](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE10114)
* Figure S4 data are from Lee et al. (2007) [@Lee:2007aa]
   - [ArrayExpress database](http://www.ebi.ac.uk/arrayexpress/) 
under accession code [E-MEXP-1172](https://www.ebi.ac.uk/biostudies/ArrayExpress/studies/E-MEXP-1172?query=E-MEXP-1172).

* Supplement the figures in the report with others
which further characterize the operating characteristics of the
proposed pre-processing step (called normalization in the paper)

<br/>

## Notes

### Normalization Verification - Minimal Requirement {-}

* "Normalization" is the processing of data in a dataset with the intent of removing
unwanted (technical) variability, while maintaining the wanted (biological) variability.

<p><p/>
* To assert the validity of a normalization procedure, at minimum the following must be 
examined and summarized, as appropriate:
   - the vector of values being normalized, before and after, and this done for studies
which have unwanted variability of the sort expected,
as well as for studies which don't seem to bear any significant
amount of unwanted variability
   - the vector of site scores (for detection, for example)
      - for known positive and known negative sites,
      - and this for studies which have unwanted variability of the sort expected
and for studies which don't.

* The validation dataset  should include typical samples with typical 
levels of unwanted variability

<br/>

### Quantile Normalization **Does Not** Require Identical Distributions {-}

<p><p/>
* In RMA, quantile normalization was used because the resulting probeset summaries
performed  best at the typical analysis tasks the microarray data were used for
(Irizarry et al. (2003) [@Irizarry:2003aa], Bolstad et al. (2003) [@Bolstad:2003aa].
   - not because this was dictated by some modeling assumptions.

<p><p/>
* Note that when working with chips, local image artifacts are part of every dataset
	and my be useful in as assessment exercise.

 ################################################################### -->

<br/>

# Spatial Transcriptomics 


* 10x  Visium spatial data consists of spots comprising 1-10 cells
   - working out what’s in a spot  one of the challenges of
working with these data.

* Cellular composition is estimated by deconvolving 10x Visium data.
Several tools are available. 
   - mapping-based methods map single cells to spatial transcriptomics (ST) data based on gene expression similarity
<span style="color: blue;">
      - NovospaRc, Tangram, CellTrek, and CytoSPACE
</span>

   - deconvolution-based methods attempt to reconstruct the ST 
observations by modeling the experimental process 
<span style="color: blue;">
      - CARD, RCTD, cell2location, DestVI, SpatialDWLS, SPOTlight, STRIDE, CellDART, Celloscope, DSTG, and Stereoscope
</span>
as a combination of single cells. 


*  Compare and contrast the methods.
   -  Marie currently uses RCTD. 

<br/>

# Technical reviews 

* Technical reviews of the state of the art for analysis techniques
in the areas of research of some PhD projects would make good
presentation topics.

## [Long read-based analysis of cancer genomes](https://www.wehi.edu.au/project/long-read-based-analysis-of-cancer-genomes/) 

* Oxford Nanopore, Pacbio and other long-read sequencing technologies
enable deep characterisation of cancer genomes, but remain relatively
inaccessible. Based around existing and planned long read data from
matched tumour and normal (blood) genomes, as well as
simulated data, this project will deliver pipelines and new methods for
long read-based characterisation of cancer genomes.


<br/>


## [Deep learning-based latent space analysis for cancer omics](https://www.wehi.edu.au/project/deep-learning-based-latent-space-analysis-for-cancer-omics/)

* Deep learning methods are being used to identify lower dimensional
latent spaces underlying complex disease datasets. Interpretation of
the latent dimensions is possible using different approaches, but
there is much room to improve understanding best practice.

* The goal of this project is to explore and refine the use
of deep learning methods to analyse and interpret latent spaces of complex
cancer omics data.

<br/>

## [Non-negative matrix factorization in cancer omics](https://www.wehi.edu.au/project/non-negative-matrix-factorization-in-cancer-omics/)

* Non-negative matrix factorisation (NMF) is an unsupervised
machine learning technique for identifying patterns in data. NMF has a
long history of use in bioinformatics and a number of NMF methods
have been developed specifically for omics data. This also includes semi-supervised approaches to NMF.

* However, how to best apply unsupervised or semi-supervised
approaches is currently not clear. This project will deliver a benchmark
of NMF methods, and develop new methods and guidelines to
make these tools more accessible and their results more interpretable.

<br/>


```{r , child=c('_index.Rmd'), eval=F}
```

<br/>

# References {#references}
   
<div id="refs"></div>
     
  
 
```{r, echo=FALSE}
  knitr::knit_exit()
```
#############################################################
#############################################################
<!--
* Comparative Illustration of the use of AI Agents 
   -  For Article Review
      - claude (v3.7 , 4.?)
      - chatGPT
      - ???
   - For literature review
-->
#############################################################
#############################################################

